---
title: "Diamonds Dataset Analysis"
author: "Ammar Al-Hawashem"
date: "28/09/2021"
output: 
  html_document:
    df_print: paged

---

## **Part1:**
### Set-up
Let's brong the built-in dataset to our environment: **data(diamonds)**

## **Part2:**
### Dataset's variabls with their descriptions:
Variable | Description
-- | -------------
price | Price in US dollars ($326 – $18,823)
carat | Weight of the diamond (0.2 – 5.01)
cut |Quality of the cut (Fair, Good, Very Good, Premium, Ideal)
color |Diamond colour, from D (best) to J (worst)
clarity |A measurement of how clear the diamond is (I1 (worst), SI2, SI1, VS2, VS1, VVS2, VVS1, IF (best))
x |Length in mm (0 – 10.74)
y |Width in mm (0 – 58.9)
z |Depth in mm (0 – 31.8)
depth |	Total depth percentage = z / mean(x, y) = 2 * z / (x + y) (43–79)
table |	Width of top of diamond relative to widest point (43–95)

## **Part3:**
### Exploring the structure of the dataset

Firstly, install tidyverse package: **library(tidyverse)** since some of its functions will be used frequently 

```{r, include=FALSE}
library(tidyverse)
```
```{r}
glimpse(diamonds) 
# OR  use the base function: str(diamonds)
```
## **Part4:**
### Summary of the dataset
To build a great intuition about the dataset, **let's use:**
```{r}
summary(diamonds)
```
## **Part5:**
### Apply some calculations

*A- How many diamonds with a clarity of category “IF” are present in the data-set?*

There are many way to calculate this. Here are four different methods:

1. The nested way:
```{r, echo = TRUE, eval = FALSE}
nrow(diamonds[diamonds$clarity=="IF",])
```
2. By using the forward piping of tidyverse with the exact name of variable (column):
```{r, echo = TRUE, eval = FALSE}
diamonds %>% 
    filter(clarity=="IF") %>% 
    nrow() -> ifnum
    ifnum
```
3. The shortest way:
```{r, echo = TRUE, eval = FALSE}
sum(diamonds$clarity=="IF")
```
4. By using the forward piping of tidyverse with the index of variable (column):
```{r, echo = TRUE, eval = FALSE}
clarity_IF <- diamonds[4] %>%
      filter(clarity == "IF")
    count(clarity_IF)
```
All of them will give us the same result: **`r sum(diamonds$clarity=="IF")`**. However, 1,2 & 3 as a vector while 4 is as tibble


*B- How many diamonds in each clarity category are present in the data-set?*
```{r}
table(diamonds$clarity) # a named vector 
```

*C- What fraction of the total do the clarity of category “IF” represent?*

Both of the following couple of methods give the solution as a vector
``` 
ifnum/nrow(diamonds) #ifnum was defined in the previous question
```
```{r}
nrow(diamonds[diamonds$clarity=="IF",])/nrow(diamonds)
```

*D- What proportion of the whole is made up of each category of clarity?*
```{r}
diamonds %>% 
  group_by(clarity) %>% 
  summarise(proportion = length(clarity)/nrow(diamonds))
                         # or n() instead of length(clarity) 
```
*E- What is the cheapest diamond price overall?*

1. To get just the cheapest price, use:
```{r}
diamonds %>%  
  summarise(cheapest = min(price))             # As a tibble
  #OR
  min(diamonds$price)                          # AS a vector
```
2. To get the full information of the cheapest diamond/s, use:
```{r}
diamonds[diamonds$price==min(diamonds$price),]
```

*F- What is the range of diamond prices?*
```{r}
diamonds %>% 
  summarise(range = max(diamonds$price) - min(diamonds$price))
```

*G- What is the average diamond price in each category of cut and color?*
```{r}
diamonds %>%
  group_by(color,cut) %>% 
  summarise(average = mean(price)) # -> avgPrice_cut_color view(avgPrice_cut_color)
```
*H- Create a 2-dimensional vector that counts the number of diamonds in each category of cut and color?*
```{r paged.print=TRUE}
table(diamonds$cut, diamonds$color) # kind of a matrix
```

## **Part6:**
### Construct some plots
1. Price Vs color 
```{r}
ggplot(diamonds, aes(x=color, y=price)) +  geom_boxplot(fill='green')
```

2. Price Vs cut 
```{r}
ggplot(diamonds, aes(x=cut, y=price)) +  geom_boxplot(fill='green')
```

3. Frequency Vs cut
```{r}
barplot(table(diamonds$cut))
```


